---
title: "Human  Activities Analysis"
author: "Ameen AboDabash"
date: "April 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Setting the Scene
####Loading and preprocessing the data

Loading required packages.

```{r  ,results='hide',message=FALSE}
packages<-c("data.table","dplyr","lattice")
sapply(packages, require, character.only = TRUE)

```

Unzip data file and load it to be processed, no need to download the files, already exsit in the forked repo :

```{r}
        
if( !dir.exists("data-raw"))
        dir.create("data-raw")

unzip("activity.zip" , exdir = "data-raw")
actData <- fread(paste("data-raw","activity.csv",sep="\\"))## %>% na.omit()

dim(actData)

```

Good to have a quick look at the data
```{r}
summary(actData)

head(actData)

```

###What is mean total number of steps taken per day?

```{r, fig.height=5}

with(aggregate(steps~date,actData, sum), hist(steps))

```

 
#####Mean and Median already know before by dispalying the summary of cleaned Dataset above, but lets write it agian:
```{r , results='markup'}

 mean(actData$steps)
 median(actData$steps)

```

###What is the average daily activity pattern?


```{r ,  fig.height=5}


with(aggregate(steps~interval,actData, mean),  plot(interval, steps,type="l",
             main = 'Average Steps by Time Interval',
             xlab = 'Time Interval',
             ylab = 'Average Number of Steps')
        )

```

#####Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
actData$interval[which.max(actData$steps)]

```



###Imputing missing values

```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
nrow(actData[is.na(steps),])


#Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
#Creating new 
actDataNComplete <- actData
actDataNComplete[is.na(actDataNComplete$steps), "steps"] <- 0


```


```{r, fig.height=5}

with(aggregate(steps~date,actDataNComplete, sum), hist(steps, main = "Hist Steps per day (NA Removed)"))

```

 
#####Mean and Median already know before by dispalying the summary of cleaned Dataset above, but lets write it agian:

```{r , results='markup'}

 mean(actDataNComplete$steps)
 median(actDataNComplete$steps)
 
```
###Are there differences in activity patterns between weekdays and weekends?

####1-Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
actDataNComplete$day=ifelse(as.POSIXlt(as.Date(actDataNComplete$date))$wday%%6==0,
                          "weekend","weekday")

# For Sunday and Saturday : weekend, Other days : weekday 
actDataNComplete$day=factor(actDataNComplete$day,levels=c("weekday","weekend"))

```



####2-Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}

with(aggregate(steps~interval+day,actDataNComplete,mean),xyplot(steps~interval|factor(day),aspect=1/2,type="l"))


```


Thanks


